---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.11.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# Excercise 1


mean: the sum of the items divided by the number of items. <br>
median: the middle item in a sorted list of items <br>
mode: the item with the most frequency <br>
variance: the average sum of the difference from the mean squared. <br>
standard deviation: the square root of the variance. <br>
histogram: a graph where bars represent the frequency of an item. <br>
Normal Distribution: a distribution that forms a bell shape where the center has the most frequency and the tails on both sides are less frequent. <br>
Poisson Distribution: a distribution where one item is more frequent than the rest of the possible outcomes. It defers from a Normal Distribution as it does not always look symmetric.


# Excercise 3

```{r}
r_datasets = data()
```

```{r}
r_datasets
```

```{r}
head(PlantGrowth)
```

```{r}
summary(PlantGrowth)
```

```{r}
library(ggplot2)
ggplot(PlantGrowth, aes(y=weight)) + geom_boxplot(aes(color = group, size=5))
```

```{r}
ggplot(PlantGrowth, aes(y=weight)) + geom_boxplot(aes(color = 'balck', size=5))
```

The dataset I choose is PlantGrowth. It contains the weight of plants from three different groups. The Average weight of all plants is 5.073 while the median is 5.155. This means that 50% of the plants weight more than 5.155, and the weight for a plant is on average around 5.073. 


# Excercise 4

```{r}
head(trees)
```

```{r}
hist(trees$Height)
```

For this exercise I selected the trees dataset. I plotted the histogram of the Height column. Here we can see how the distribution looks normal as most trees fall in the 70-80 height while fewer fall on the tail ends creating a bell curve.
