---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.11.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# Exercise 2

```{r}
?ChickWeight
```

```{r}
str(ChickWeight)
```

```{r}
summary(ChickWeight)
```

The four variables / columns are weight, Time, Chick and Diet.

```{r}
dim(ChickWeight)
```

The first number in the dim(ChickWeight) command referes to the number of rows in the ChickWeight dataframe. This means that there is a total of 578 records done in the experiment. There are 578 unique rows where each row has a unique combinations of a chick with a specific diet at one time interval. 


# Exercise 3

```{r}
summary(ChickWeight$weight)
```

The summary command gives us the high level statistics of the weight column in the ChickWeight dataframe.

```{r}
head(ChickWeight$weight)
```

The head command gives us the first 5 values of the weight column in the ChickWeight dataframe.

```{r}
mean(ChickWeight$weight)
```

The mean command gives us the high average of the weight column in the ChickWeight dataframe.

```{r}
myChkWts <- ChickWeight$weight
```

The command above saves the weight column from the ChickWeight into a variable called myChkWts

```{r}
quantile(myChkWts, 0.50)
```

The quantile command above returns the 50% quantile (the median) of the myChkWts variable. I.e., it returns the median weight of the chicks dataframe.


# Exercise 4

```{r}
myChkWts <- ChickWeight$weight
```

```{r}
quantile(myChkWts, .025)
quantile(myChkWts, .975)
```

```{r}
summary(myChkWts)
```

```{r}
hist(myChkWts)
abline(v=quantile(myChkWts, .975), col="red")
abline(v=quantile(myChkWts, .025), col="red")
abline(v=mean(myChkWts), col="green")
abline(v=quantile(myChkWts, .5), col="red")
```

The variable myChWts looks like a poisson distribution as most of the weights fall into the 50-100 bucket, and it has a much longer tail on the right side of the median. The median is 103 which means 50% of the weights lie before 103. The mean is 121.8 which means on average the weight of a chick is about 121.8. The 2.5% quantile is 41 and the 97.5% quatile is 294.575 which means that 95% of the data lies between these two points.


# Exercise 5

```{r}
chkWtsSamMeanDist <- replicate(n=1000, mean(sample(myChkWts, size=11, replace=TRUE)))
```

```{r}
hist(chkWtsSamMeanDist)
abline(v=quantile(chkWtsSamMeanDist, .975), col="red")
abline(v=quantile(chkWtsSamMeanDist, .025), col="red")
```

# Exercise 6


The main difference between a distribution of raw data and the distribution of a sampling of means is that the distribution of raw data will not always be a normal distribution while the sampling of means will always be a normal distribution as the number of means increases. The sampling of means returns a description of how the mean of the raw data is distributed.


The 2.5% and the 97.5% quantiles are very different from each other as the sampling of means centers around the raw input mean while the raw data quantiles center around the median of the data.


# Exercise 7

```{r}
chkWtsSamMeanDist_n100 <- replicate(n=1000, mean(sample(myChkWts, size=100, replace=TRUE)))
```

```{r}
mean(chkWtsSamMeanDist)
```

```{r}
mean(chkWtsSamMeanDist_n100)
```

```{r tags=c()}
hist(chkWtsSamMeanDist_n100)
abline(v=quantile(chkWtsSamMeanDist_n100, .975), col="red")
abline(v=quantile(chkWtsSamMeanDist_n100, .025), col="red")
```

Even though both of the distributions have a center close to 120, the sampling distribution with n=100 has a higher confidence since the sample size is bigger. This means that the 2.5% and 97.5% quantiles will be more tightly bound.
